Evolution of precopulatory and postcopulatory inbreeding behaviour

### 2016-09-13 ms_pre_post.Rmd
Add some empirical studies to evolution of polyandry to avoid inbreeding

@@

\vspace{60 mm}

>>>>

Inbred offspring are typically less viable than outbred offspring, and when such inbreeding depression is strong, it is predicted to cause selection for inbreeding avoidance [Duthie and Reid *in press*; @Parker1979; @Parker2006; @Pusey1996; @Kokko2006]. In animals, inbreeding avoidance can be realised through both pre-copulatory and post-copulatory mechanisms [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, the conditions under which inbreeding avoidance is predicted to evolve via pre-copulatory, post-copulatory, or both pre- and post-copulatory mechanisms has not been modelled. There is therefore no framework for predicting what mechanisms of inbreeding avoidance are most likely to evolve in empirical populations where risk of fitness loss caused by inbreeding is high. Importantly, evolution of post-copulatory mechanisms of inbreeding avoidance requires females to mate multiply, meaning that post-copulatory inbreeding avoidance cannot occur unless polyandry has evolved in a population. Accordingly, we modelled evolution of three separate traits affecting polyandry and inbreeding avoidance, including tendency to engage in polyandry, inbreeding mating strategy (pre-copulatory), and inbreeding fertilisation strategy (post-copulatory). Using *in silico* experiments, we tested four specific hypothesis to predict evolution of tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy given diverse costs and evolutionary constraints. We found that (1) selection for polyandry was greatly strengthened given concurrent evolution of inbreeding fertilisation strategy, (2) selection was strong for low cost, and weak but not negligible for high cost, mechanisms of inbreeding avoidance, (3) constraints on evolution of pre-copulatory inbreeding avoidance greatly strengthened selection for post-copulatory inbreeding avoidance, but not vice versa, and (4) pre-existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of post-copulatory inbreeding avoidance, but not vice versa. Our results therefore provide a novel framework for approaching empirical studies of inbreeding avoidance, predicting evolution of polyandry and inbreeding avoidance through pre- and post-copulatory mechanisms.

<!--- (1) selection for polyandry is strengthened given adaptive inbreeding avoidance through fertilisation bias, (2) selection for pre- versus post-copulatory inbreeding avoidance will be strong for the least costly, and negligible for the most costly, mechanism of inbreeding avoidance, (3) selection for pre- or post-copulatory inbreeding avoidance will be strengthened if evolutionary constraints preclude inbreeding avoidance in the complementary mechanism, and (4) selection for pre- versus post-copulatory inbreeding avoidance will be weakened if a complementary adaptive mechanism of inbreeding avoidance is already fixed in a population.--->

Evolution of polyandry to avoid inbreeding
----------------------------------------------

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. Recently, @Duthie found the conditions under which selection for polyandry to avoid inbreeding is predicted to be highly restrictive, requiring strong constraints on the availability of initial compared to additional mates, or the conditional expression of polyandry such that females only engaged in polyandry when their initial choice of mate was of low quality. However, @Duthie assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express choice after mating had occurred. In contrast, we found that when females could express choice after mating through biased fertilisation, selection for polyandry was greatly strengthened (Figure 1A), especially when constraints restricted evolution of inbreeding avoidance through mate choice (Figure 3). The proposition that polyandry might facilitate female choice among males of varying compatability is not new [e.g., @Zeh1997], but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for empirical studies of evolution of polyandry.

We predict evolution of polyandry in populations where inbreeding depression is severe and evolution of inbreeding avoidance through post-copulatory mechanisms is not constrained. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g. @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. For example, @Pizzari2004 find that the number of sperm arriving to eggs is lower for full siblings of female red jungefowl (*Gallus gallus*) than it is for non-relatives, even after controlling for effects of order of oviposition and social familiarity. Similarly, Firman and Simmons [-@Firman2008; -@Firman2015] have shown evidence of post-copulatory inbreeding avoidance in house mice, finding fertilisation to be biased toward unrelated males independent of the order in which mating occurs [@Firman2008] most likely due to post-copulatory mechanisms expressed through the secretion of gametic proteins originating in female's eggs [@Firman2015]. Interestingly, female house mice have also been shown to express pre-copulatory inbreeding avoidance  [@Potts1991; @Roberts2003]. Our results suggest that post-copulatory inbreeding avoidance is unlikely to evolve when females can already avoid inbreeding through pre-copulatory mechanisms (Figure 5), meaning that evolution of post-copulatory inbreeding avoidance could reasonably be hypothesised to have preceded evolution of pre-copulatory inbreeding avoidance in mice.  <!--- start talking abou these now --->

<!--- Olsson1996; Pizzari2004; Firman2008; Bretman2009; Gasparini2011; Tuni2013; Firman2015 --->

@@

\vspace{150 mm}
